---
title: "Themes"
date: "2026-05-14"
date_end: "2026-05-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 8,
  fig.height = 4.8,
  fig.align = "center",
  collapse = TRUE,
  dev = "png"
)
set.seed(1234)

library(showtext)
font_add_google("Roboto Condensed", "Roboto Condensed")
showtext_auto()
```

The [lesson for this week's session](/lesson/05-lesson.qmd) is a fairly comprehensive introduction to using the `theme()` function in ggplot, and [this page by Henry Wang](https://henrywang.nl/ggplot2-theme-elements-demonstration/) is a good cheat sheet for remembering which theme elements are which on a plot—and I like [this PDF cheatsheet by Clara Granell](https://github.com/claragranell/ggplot2/blob/main/ggplot_theme_system_cheatsheet.pdf) even better.

For [your exercise](/assignment/05-exercise.qmd), you're going to create the world's ugliest plot. For this example, we'll use the principles of CRAP to make a great theme.

I'm going to build the theme semi-incrementally here. Instead of showing how the plot updates with each change in setting, I do most of the updates all at once, with tons of comments explaining what each line does. **Importantly**, I did *not* write this all at once. When you're tinkering with themes, you generally start with something like `theme_minimal()` or `theme_bw()` and then gradually add new things to `theme()`, like modifying `plot.title`, then `plot.subtitle`, etc. It's a very iterative process with lots of tinkering. Because of this, **there is no live-coding video for this example**—it would be incredibly long and boring. Instead, look through each of the lines and see what they're doing.

For this example, I'm going to use the `gapminder` dataset that we've been using throughout this week. Instead of using the CSV file like we did before, we'll load the data from the {gapminder} package. Once you run `library(gapminder)`, you'll automatically have access to a dataset named `gapminder`.

We'll also use the [Roboto Condensed font](https://fonts.google.com/specimen/Roboto+Condensed) in the theme. The setup chunk at the top of this document loads it automatically via the `{showtext}` package, so no manual font installation is needed.


## Bonus: {ggThemeAssist}

Before we dive into building a theme from scratch, let's look at a tool that can help you explore theme elements interactively. {ggThemeAssist} provides a point-and-click interface for tweaking theme settings, and then generates the corresponding code for you — great for discovering what's possible before you start writing theme code by hand.

Install it with:

```{r install-ggthemeassist, eval=FALSE}
remotes::install_github("timmarchand/ggthemeassist")
library(ggThemeAssist)
```

To use it, first create a ggplot object, select its name in your script, then go to **Addins → ggplot Theme Assistant**. Here's a simple plot to try it on:

```{r ggthemeassist-demo, eval = FALSE}
library(tidyverse)

demo_plot <- ggplot(data = mpg,
                    mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point(size = 3) +
  labs(x = "Displacement", y = "Highway MPG", color = "Car class",
       title = "Heavier cars get worse mileage",
       subtitle = "Except for two-seaters?",
       caption = "Source: ggplot2::mpg")

demo_plot
```

Select `demo_plot` in your script and launch the addin. Once you're happy with the result, the generated `theme()` code gets inserted directly into your script. Here's a brief video showing the workflow:

<div class="ratio ratio-16x9">
<iframe src="https://www.youtube.com/embed/9ldrTCUSReM" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen="" frameborder="0"></iframe>
</div>


## Basic plot

When I'm creating a theme, I like to use a basic plot with everything that might show up, complete with a title, subtitle, caption, legend, facets, and other elements.

```{r basic-plot, warning=FALSE, message=FALSE}
library(tidyverse)  # For ggplot, dplyr, and friends
library(gapminder)  # For gapminder data
library(scales)     # For nice axis labels

gapminder_filtered <- gapminder |> 
  filter(year > 2000)

base_plot <- ggplot(data = gapminder_filtered,
                    mapping = aes(x = gdpPercap, y = lifeExp, 
                                  color = continent, size = pop)) +
  geom_point() +
  # Use dollars, and get rid of the cents part (i.e. $300 instead of $300.00)
  scale_x_log10(labels = label_dollar(accuracy = 1)) +
  # Format with commas
  scale_size_continuous(labels = label_comma()) +
  # Use viridis
  scale_color_viridis_d(option = "plasma", end = 0.9) +
  labs(x = "GDP per capita", y = "Life expectancy",
       color = "Continent", size = "Population",
       title = "Here's a cool title",
       subtitle = "And here's a neat subtitle",
       caption = "Source: The Gapminder Project") +
  facet_wrap(vars(year))

base_plot
```

Now we have `base_plot` to work with. Here's what it looks like with `theme_minimal()` applied to it:

```{r base-minimal}
base_plot +
  theme_minimal()
```

That gets rid of the grey background and is a good start, but we can make lots of improvements. First let's deal with the gridlines. There are too many. We can get rid of the minor gridlines by setting them to `element_blank()`:

```{r theme1}
base_plot +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
```

Next let's add some typographic contrast. We'll use Roboto Condensed Regular as the base font, which is already loaded via `{showtext}` in the setup chunk.

We'll use the font as the `base_family` argument. Note how I make it bold with `face` and change the size with `rel()`. Instead of manually setting some arbitrary size, I use `rel()` to resize the text in relation to the `base_size` argument. Using `rel(1.7)` means 1.7 × `base_size`, or 20.4. That will rescale according to whatever `base_size` is—if I shrink it to `base_size = 8`, the title will scale down accordingly.

```{r theme2}
plot_with_good_typography <- base_plot +
  theme_minimal(base_family = "Roboto Condensed", base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        # Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Italic, smaller, grey caption that is left-aligned
        plot.caption = element_text(face = "italic", size = rel(0.7), 
                                    color = "grey70", hjust = 0),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold, slightly larger facet titles that are left-aligned for the sake of repetition
        strip.text = element_text(face = "bold", size = rel(1.1), hjust = 0),
        # Bold axis titles
        axis.title = element_text(face = "bold"),
        # Add some space above the x-axis title and make it left-aligned
        axis.title.x = element_text(margin = margin(t = 10), hjust = 0),
        # Add some space to the right of the y-axis title and make it top-aligned
        axis.title.y = element_text(margin = margin(r = 10), hjust = 1))
plot_with_good_typography
```

Whoa. That gets us most of the way there! We have good contrast with the typography, with the strong bold and the lighter regular font (**✓ contrast**). Everything is aligned left (**✓ alignment** and **✓ repetition**). By moving the axis titles a little bit away from the labels, we've enhanced proximity, since they were too close together (**✓ proximity**). We repeat grey in both the caption and the subtitle (**✓ repetition**).

The only thing I don't like is that the 2002 facet label isn't quite aligned with the title and subtitle. This is because the facet labels are in boxes along the top of each plot, and in some themes (like `theme_grey()` and `theme_bw()`) those facet labels have grey backgrounds. We can add a background, which will then be perfectly aligned with the title and subtitle.

```{r theme3}
plot_with_good_typography +
  # Add a light grey background to the facet titles, with no borders
  theme(strip.background = element_rect(fill = "grey90", color = NA),
        # Add a thin grey border around all the plots to tie in the facet titles
        panel.border = element_rect(color = "grey90", fill = NA))
```

`r emoji::emoji("chef")` `r emoji::emoji("kiss")`! That looks great!

To save ourselves time in the future, we can store this whole thing as an object that we can then reuse on other plots:

```{r theme-store}
my_pretty_theme <- theme_minimal(base_family = "Roboto Condensed", base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        # Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Italic, smaller, grey caption that is left-aligned
        plot.caption = element_text(face = "italic", size = rel(0.7), 
                                    color = "grey70", hjust = 0),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold, slightly larger facet titles that are left-aligned for the sake of repetition
        strip.text = element_text(face = "bold", size = rel(1.1), hjust = 0),
        # Bold axis titles
        axis.title = element_text(face = "bold"),
        # Add some space above the x-axis title and make it left-aligned
        axis.title.x = element_text(margin = margin(t = 10), hjust = 0),
        # Add some space to the right of the y-axis title and make it top-aligned
        axis.title.y = element_text(margin = margin(r = 10), hjust = 1),
        # Add a light grey background to the facet titles, with no borders
        strip.background = element_rect(fill = "grey90", color = NA),
        # Add a thin grey border around all the plots to tie in the facet titles
        panel.border = element_rect(color = "grey90", fill = NA))
```

Now we can use it on any plot. Remember that first plot you made in your exercise from session 1 with the `cars` dataset? Let's throw this theme on it! (only here the dataset is named `mpg` instead of `cars`; the `mpg` dataset is loaded invisibly whenever you load ggplot)

```{r mpg-example}
mpg_example <- ggplot(data = mpg, 
                      mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point(size = 3) + 
  scale_color_viridis_d() +
  facet_wrap(vars(drv)) +
  labs(x = "Displacement", y = "Highway MPG", color = "Car class",
       title = "Heavier cars get worse mileage",
       subtitle = "Except two-seaters?",
       caption = "Here's a caption") +
  my_pretty_theme

mpg_example
```

Super neat!


## Nice pre-built themes

This custom theme we just made is just one iteration of a theme. There are countless ways to tinker with a theme and have it meet the different CRAP principles. People have even published their own themes in different R packages. Check these out to see lots of different examples:

- [{hrbrthemes}](https://github.com/hrbrmstr/hrbrthemes)
- [{ggthemes}](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)
- [{ggthemr}](https://github.com/cttobin/ggthemr)
- [{ggtech}](https://github.com/ricardo-bion/ggtech)
- [{tvthemes}](https://ryo-n7.github.io/2019-05-16-introducing-tvthemes-package/)
- [{ggpomological}](https://www.garrickadenbuie.com/project/ggpomological/) (this one is incredible!)

Check [this blog post](https://rfortherestofus.com/2019/08/themes-to-improve-your-ggplot-figures/) for examples of a bunch of others.


## Saving plots

If we want to save these plots, we can use `ggsave()`. For that to work, we need to store the plot as an object, which I already did in the examples above:

```{r store-plot-example, eval=FALSE}
name_of_plot_object <- ggplot(...)
```

We then feed our saved plot object to `ggsave()` and specify the filename and dimensions we want to use. If we're using PNG, we don't need to worry about any extra options. If we're using PDF, [we need to tell R to use the Cairo PDF writing engine](https://www.andrewheiss.com/blog/2017/09/27/working-with-r-cairo-graphics-custom-fonts-and-ggplot/) instead of R's normal one, since R's normal one can't deal with custom fonts.

```{r save-plots, eval=FALSE}
# Add my_pretty_theme to the gapminder base_plot and save as an object
final_gapminder_plot <- base_plot +
  my_pretty_theme

# Save as PNG and PDF
ggsave("fancy_gapminder.png", final_gapminder_plot,
       width = 8, height = 5, units = "in")
ggsave("fancy_gapminder.pdf", final_gapminder_plot,
       width = 8, height = 5, units = "in", device = cairo_pdf)

# Save the mpg plot as PNG and PDF
ggsave("fancy_mpg.png", mpg_example,
       width = 8, height = 5, units = "in")
ggsave("fancy_mpg.pdf", mpg_example,
       width = 8, height = 5, units = "in", device = cairo_pdf)
```
